\documentclass[12pt,14paper]{scrartcl}
\usepackage{listings,UoYstyle}
\lstset{language=[latex]tex,basicstyle=\sffamily}
\title{\LaTeXe\ Problems for CSW}
\author{Jeremy Jacob}
\date{2008 October 16}
\begin{document}
\maketitle

\section{Problems}

These exercises are designed to be run using the \TeX{}live
distribution.  It is best to use \lstinline{pdflatex} (rather than
\lstinline{latex}), and view the output using a PDF viewer.

\begin{enumerate}
\item Create an A4 document of the class \lstinline{article} whose
  title is `Greetings', that has you as the author and today's date.
  The body should have two sections, titled `Welcome' and `Farewell',
  in that order.  Each section should have two sentences.  The first
  should be one of `Hello world' or `Goodbye world' as appropriate.
  The second should be a sentence of the shape `See also Section~n',
  where `n' is the number of the other section.  The numbers should
  not appear in your source [hint: use symbolic labels and
  references].
\item Create an identical document, but make it of class
  \lstinline{scrartcl}.  Do you notice a difference in the output?
  Which do you prefer?
\item In the document you have created so far, you should have typed
  `See also' twice.  Introduce a macro, \lstinline{\sa}, to typeset
  `See also', and use it in both places (do not forget to protect the
  following space).
\item In the document you have created so far, you should have typed
  the structure `\textless salute\textgreater world.' twice, once for
  salutation `Hello' and once for salutation `Goodbye'.  Introduce a
  macro, \lstinline{\sw}, with one parameter that is the salutation,
  and use it in both places.
\item If the package \lstinline{hyperref} is loaded\footnote{As
    \lstinline{UoYCSproject} does.} then, as well as automatically
  generating bookmarks and internal links, the macro
  \lstinline|\autoref{label}| is available.  This macro knows the kind
  of reference that a label refers to, so you do not have to type
  `\lstinline|Section~|', and so on.  Make your document load the
  package, and modify your document to use \lstinline{\autoref} rather
  than \lstinline{\ref}.
\item Your document consists of one structure repeated twice.  The
  structure has three variables.  In the first instance it is
  instantiated with \lstinline{Welcome}, \lstinline{Hello} and
  \lstinline{Farewell}; in the second it is instantiated with
  \lstinline{Farewell}, \lstinline{Goodbye} and \lstinline{Welcome}.
  Create a 3-parameter macro, \lstinline{\st}, that captures this
  structure and use it.  The body of your document should be
  \lstinline{\maketitle} followed by two calls to your new macro, and
  nothing else.\footnote{If you wish you can add a package to change
    fonts from the default set; to create a good PDF document this
    should always be done.  Suitable packages are \lstinline{mathpazo}
    (Palatino) and \lstinline{mathptmx} (Times Roman), both combined
    with \lstinline{helvet,courier}.  The package
    \lstinline{microtype} is also recommended.}
\end{enumerate}

\newpage

\section{Possible solutions}

\begin{enumerate}
\item 
  This assumes I wrote it today.  Your answer should have a different
  name and date\ldots\ It may also have different labels.
  \begin{lstlisting}
    \documentclass[a4paper]{article}
    \title{Greetings}
    \author{Jeremy Jacob}
    \begin{document}
    \maketitle
    \section{Welcome}\label{sec:Welcome}
    Hello world.  See also Section~\ref{sec:Farewell}.
    \section{Farewell}\label{sec:Farewell}
    Goodbye world.  See also Section~\ref{sec:Welcome}.
    \end{document}
  \end{lstlisting}
\item For completeness:
  \begin{lstlisting}
    \documentclass[a4paper]{scrartcl}
    \title{Greetings}
    \author{Jeremy Jacob}
    \begin{document}
    \maketitle
    \section{Welcome}\label{sec:Welcome}
    Hello world.  See also Section~\ref{sec:Farewell}.
    \section{Farewell}\label{sec:Farewell}
    Goodbye world.  See also Section~\ref{sec:Welcome}.
    \end{document}
  \end{lstlisting}
\item\
  \begin{lstlisting}
    \documentclass[a4paper]{article}
    \title{Greetings}
    \author{Jeremy Jacob}
    \newcommand{\sa}{See also}
    \maketitle
    \begin{document}
    \section{Welcome}\label{sec:Welcome}
    Hello world.  \sa\ Section~\ref{sec:Farewell}.
    \section{Farewell}\label{sec:Farewell}
    Goodbye world.  \sa\ Section~\ref{sec:Welcome}.
    \end{document}    
  \end{lstlisting}
  \newpage
\item\
  \begin{lstlisting}
    \documentclass[a4paper]{article}
    \title{Greetings}
    \author{Jeremy Jacob}
    \newcommand{\sa}{See also}
    \newcommand{\sw}[1]{#1 world.}
    \begin{document}
    \maketitle
    \section{Welcome}\label{sec:Welcome}
    \sw{Hello}  \sa\ Section~\ref{sec:Farewell}.
    \section{Farewell}\label{sec:Farewell}
    \sw{Goodbye}  \sa\ Section~\ref{sec:Welcome}.
    \end{document}    
  \end{lstlisting}
\item\
  \begin{lstlisting}
    \documentclass[a4paper]{article}
    \usepackage{hyperref}
    \title{Greetings}
    \author{Jeremy Jacob}
    \newcommand{\sa}{See also}
    \newcommand{\sw}[1]{#1 world.}
    \begin{document}
    \maketitle
    \section{Welcome}\label{sec:Welcome}
    \sw{Hello}  \sa\ \autoref{sec:Farewell}.
    \section{Farewell}\label{sec:Farewell}
    \sw{Goodbye}  \sa\ \autoref{sec:Welcome}.
    \end{document}    
  \end{lstlisting}
  \newpage
\item\
  \begin{lstlisting}
    \documentclass[a4paper]{article}
    \usepackage{hyperref}
    \title{Greetings}
    \author{Jeremy Jacob}
    \newcommand{\sa}{See also}
    \newcommand{\sw}[1]{#1 world.}
    \newcommand{\st}[3]{%
      \section{#1}\label{sec:#1}\sw{#2}  \sa\ \autoref{sec:#3}.%
    }
    \begin{document}
    \maketitle
    \st{Welcome}{Hello}{Farewell}
    \st{Farewell}{Goodbye}{Welcome}
    \end{document}    
  \end{lstlisting}
\end{enumerate}

\end{document}
